---
title: "INF2190_Final_Methods_Code"
date: "2022-11-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(aida)
library(tinytex)
library(knitr)
```

# Loading and cleaning data
Data has been cleaned according to the process described in Michael 
Franke's Introduction to Data Analysis, Appendix D.1. However, the clean data
available in the package does not include two variables that we want to incorporate
into our model: age and education. 

```{r, include=FALSE}
mc_data_raw <- aida::data_MC_raw

block_levels <- c("reaction", "goNoGo", "discrimination") # ordering of blocks for plotting, etc. 

mc_data_preprocessed <- mc_data_raw %>% 
  separate(trial_type, c("block", "stage"), sep = "_", remove = FALSE) %>%
  mutate(comments = ifelse(is.na(comments), "non given", comments)) %>% 
  filter(stage == "main") %>% 
  mutate(
    block = factor(block, ordered = T, levels = block_levels),
    response = ifelse(is.na(response), "none", response)
  ) %>%
  filter(response != "wait") %>% 
  rename(
    handedness = languages, # variable name is simply wrong
    total_time_spent = timeSpent
  ) %>% 
  select(
    submission_id, 
    trial_number, 
    block, 
    stimulus, 
    RT, 
    handedness, 
    gender, 
    age, 
    education, 
    total_time_spent,
    comments
  )

# summary stats (means) for participants
d_sum_stats_participants <- mc_data_preprocessed %>% 
  group_by(submission_id, block) %>% 
  summarise(
    mean_P = mean(RT)
  )

# summary stats (means and SDs) for conditions
d_sum_stats_conditions <- mc_data_preprocessed %>% 
  group_by(block) %>% 
  summarise(
    mean_C = mean(RT),
    sd_C   = sd(RT)
  )
  
d_sum_stats_participants <- 
  full_join(
    d_sum_stats_participants,
    d_sum_stats_conditions,
    by = "block"
  ) %>% 
  mutate(
    outlier_P = abs(mean_P - mean_C) > 2 * sd_C
  )

mc_data_cleaned <- mc_data_preprocessed %>% 
  filter(submission_id != d_sum_stats_participants$submission_id[1])

mc_data_cleaned <- mc_data_cleaned %>% 
  full_join(
    d_sum_stats_conditions,
    by = "block"
  ) %>% 
  mutate(
    trial_type = case_when(
      abs(RT - mean_C) > 2 * sd_C ~ "too far from mean",
      RT < 100 ~ "< 100ms",
      TRUE ~ "acceptable"
    ) %>% factor(levels = c("acceptable", "< 100ms", "too far from mean")),
    trial = 1:nrow(mc_data_cleaned)
  )

mc_data_cleaned <- mc_data_cleaned %>% 
  filter(trial_type == "acceptable")

write.csv(mc_data_cleaned, file = "clean_data.csv")
```

## Coding categorical variables as numeric ones
To make analysis easier, binary variables of interest should be coded as numeric
instead of categorical variables. 

```{r}
mc_data_cleaned <- mc_data_cleaned %>%
  mutate(education_num = ifelse(education == "high school / college", 0, 1), 
         handedness_num = ifelse(handedness == "left" | handedness == "Left", 0, 1), 
         gender_num = ifelse(gender == "male", 0, 1))
```

Previewing the data: 
```{r, echo=FALSE}
kable(head(mc_data_cleaned))
```

# Correlation matrix
There are two models we can construct: one with the dependent variable being RT 
and another with the dependent variable being total_time_spent. First, we pull 
all of the characteristics we were interested in analyzing and compare their 
individual correlations with RT and total_time_spent. The independent variables 
of interest are: handedness, gender, age, and education. 

Table 1: Correlations of each independent variable with RT. 
```{r, echo=FALSE}
RT_cor_matrix <- mc_data_cleaned %>%
  summarize(RT_handedness_cor = cor(RT, handedness_num, use = "complete.obs"), 
            RT_gender_cor = cor(RT, gender_num), 
            RT_age_cor = cor(RT, age), 
            RT_education_cor = cor(RT, education_num, use = "complete.obs"))
kable(RT_cor_matrix)
```

Table 2: Correlations of each independent variable with total_time_spent. 
```{r, echo=FALSE}
total_time_spent_cor_matrix <- mc_data_cleaned %>%
  summarize(time_handedness_cor = cor(total_time_spent, handedness_num, use = "complete.obs"), 
            time_gender_cor = cor(total_time_spent, gender_num), 
            time_age_cor = cor(total_time_spent, age), 
            time_education_cor = cor(total_time_spent, education_num, use = "complete.obs"))
kable(total_time_spent_cor_matrix)
```

# Hypothesis testing

## Model 1: RT on demographic characteristics
We can construct a multiple linear regression of RT on each independent variable. 
To assess the significance of each independent variable, we assume $\alpha$ = 0.05. 
```{r}
RT_lm_model <- lm(RT ~ handedness_num + gender_num + age + education_num, 
                    data = mc_data_cleaned)
kable(summary(RT_lm_model)$coef)
```

Handedness, age, and education all have a statistically significant relationship
with reaction time. 

## Model 2: total_time_spent on demographic characteristics
We can construct a multiple linear regression of total_time_spent on each independent variable. 
To assess the significance of each independent variable, we assume $\alpha$ = 0.05. 
```{r}
time_lm_model <- lm(total_time_spent ~ handedness_num + gender_num + age + education_num, 
                    data = mc_data_cleaned)
kable(summary(time_lm_model)$coef)
```

Handedness, gender, and age all have a statistically significant relationship with total 
time spent on all 3 reaction time tasks in the study. 

# Plots
To come. 